---
title: "BIOL432_Assignment1"
author: "Autumn Hodgins"
date: "2025-01-08"
output: html_document
---
#BIOL 432 Assignment 1

This assignment can be found at my [GitHub repository](https://github.com/autumnhodgins/BIOL432_Assignment1).

```{r}
#Running dataGenerato.R to make a data frame with measurments of pigeon species.
source("dataGenerato.R")

#Running volumeEstimato.R to update the data frame with volume calculations.
source("volumeEstimato.R")
```

Now I will load the data and sort it.

```{r}
#Load in the data from measurements.csv
data <- read.csv("measurements.csv")

#I will use head to check if the data looks correct
head(data)

#Load the dplyr library
library(dplyr)

#Sort the data, first by species, then by observer, then by limb volume
data <- data %>% arrange(Organism, Observer, Volume)

#I will use head to check if the data looks correct
head(data)
```

Next, I will make a table showing the average limb volume for each species.

```{r}
#Calculating the average limb volume for each species
avg_volume <- data %>% group_by(Organism) %>% summarise(Average_Volume = mean(Volume))

#Loading the library knitr so that I can make a nice looking table
library(knitr)

#Making a table to display average limb volme by species
kable(avg_volume, caption = "Average Volume by Species")
```
I will now make a table to display how many times each observer measured each species of bird.

```{r}
#Calculating how many entries exist for each observer with each species.
observation_frequency <- data %>% group_by(Organism, Observer) %>% summarise(Observations = n())

#Making a table to display observation frequencies.
kable(observation_frequency, caption = "Observation Frequency")

```
I will now make a boxplot to display the distributions of limb volumes for each species of pigeon.

```{r}
#loading in ggplot2 for making the graphs
library(ggplot2)

ggplot(data, aes(x = Organism, y = Volume)) +
  geom_boxplot() +
  theme_light() +
  labs(title = "Average Limb Volume by Species of Pigeon",
       x = "Species of Pigeon",
       y = bquote("Limb Volume (" ~ cm^3 ~ ")"))
```
Now I will make a multi-panel plot to show frequency histograms of the distribution of limb volume for each species.

```{r}
#Changing the limb volumes of my data to be in long data format so that they can be used for a multi-plot
volume_long_data <- data %>% pivot_longer(cols = Volume,
               names_to = "Measurement",
               values_to = "Value")

#Making a multi-panel plot to show the distributions of limb volume by species.
ggplot(volume_long_data, aes(x = Value)) +
  geom_histogram() +
  facet_wrap(~Organism) +
  theme_light() +
  labs(title = "Frequency Distributions of Limb Volume of Pigeon Species",
       x = bquote("Limb Volume (" ~ cm^3 ~ ")"),
       y = "Frequency")
```


